Data Specialization for a Staged Lambda Calculus

### 2015-02-13 paper/splitting.tex
work on section 4.0

In this section, we provide a {\em splitting algorithm} to translate the multistage terms of \lang\,
into equivalent but simpler forms where the stage \bbone\ and stage \bbtwo\ portions 
of the computation are {\em syntactically separated} from each other.

More precisely, for any $\coltwo e A$ which reduces to a residual $q$ (via $\diatwo e q$),
we can translate $e$ into some syntactically separated program $\mathcal{P}$ 
(via the new relation $e \splittwosym \mathcal{P}$), which
can also be reduced to the residual $q$, up to an equivalence.
We enforce the ``syntactically separated'' condition structurally by saying that 
$\mathcal{P}$ must have the form $\pipeS p l r$ where  
$p$ (called the {\em precomputation}) is a monostage term encoding all the stage \bbone\ computation that was in $e$, 
and $l.r$ (called the {\em resumer}) is a monostage term encoding all the stage \bbtwo\ computation that was in $e$.  
We can then residualize $\mathcal{P}$ with the $\sepredtwosym$ judgement, which is defined by the single rule
which evaluates the precomputation and plugs it in for $l$:
\[
\infer{\sepredtwo {\pipeS p l r} {[b/l]r}} {\reduce p b}
\]
A summary of these operations and the relationship between them is provided on the right side of \ref{fig:splittingSummary}.

>>>>

In this section, we provide a {\em splitting algorithm} to translate the multistage terms of \lang\
into a simpler but equivalent form.
The output of splitting is simpler because it is {\em syntactically separated}; that is, 
stage \bbone\ and stage \bbtwo\ are represented by distinct subterms and are only connected at the top syntactic level.
The output of splitting is equivalent (to its input) because it can be evaluated to the same residual 
that would be produced by the semantics of \ref{sec:semantics}.

More precisely, for any $\coltwo e A$ which reduces to a residual $q$ (via $\diatwo e q$),
splitting translates $e$ into some syntactically separated program $\mathcal{P}$ 
(via the new relation $e \splittwosym \mathcal{P}$), which
can also be reduced to $q$ (via the new relation $\mathcal{P} \sepredtwosym q$).
We enforce the ``syntactically separated'' condition by saying that $\mathcal{P}$ 
can only have one form, namely $\pipeS p l r$, where  
$p$ (called the {\em precomputation}) is a monostage term encoding all the stage \bbone\ computation that was in $e$, 
and $l.r$ (called the {\em resumer}) is a monostage term encoding all the stage \bbtwo\ computation that was in $e$.  
Likewise, $\sepredtwosym$ is defined by a single rule, which evaluates the precomputation and plugs it in for $l$:
\[
\infer{\sepredtwo {\pipeS p l r} {[b/l]r}} {\reduce p b}
\]
Here we call $b$ the {\em boundary value}, as it represents the communication at the boundary between stages.
A summary of these operations and the relationship between them is provided on the right side of \ref{fig:splittingSummary}.
